---
title: 'Session 9: Geospatial Data in R'
author: "Norika and Marina"
date: "2023-03-02"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1:Map Map Map!

## Preparation

This code is generating **maps of the United States and California** with marked cities. Let's load the 'maps' package and use 'us.cities' data set

```{r maps}
library(maps)
data(us.cities)
head(us.cities)
```

### Questions 1:
Explain what kind of information does us.cities dataframe have. Which columns represent the geographical information of each city?

EXTRA CELL TO ANSWER

## Drawing a map: USA

Next, it draws a map of the entire USA with the map() function. It then creates a subset of the 'us.cities' data set, selecting only those cities that are state capitals.

```{r usa }
map(database = "usa") 
capitals <- subset(us.cities, capital == 2) # subset state capitals

```


## Decorations! 
Let's add points to the map using the points() function, with the size of the points proportional to the city's population. Also, let's add a title to the map using the title() function.

```{r usa2 }
map(database = "usa") 
capitals <- subset(us.cities, capital == 2) # subset state capitals

## add points proportional to population using latitude and longitude
points(x = capitals$long, y = capitals$lat,
       cex = capitals$pop / 500000, pch = 19) 
title("US state capitals") # add a title
```

Let's try the same thing at state level!

## Drawing a map: CA

```{r ca }
map(database = "state", regions = "California")
```

In the cell below, we creates a subset of the 'us.cities' data set with only the cities in California. It then orders the California cities by population, selects the top seven, and adds points to the California map with the points() function.

```{r ca2 }
cal.cities <- subset(us.cities, subset = (country.etc == "CA"))
sind <- order(cal.cities$pop, decreasing = TRUE) # order by population
top7 <- sind[1:7] # seven cities with largest population

map(database = "state", regions = "California")

points(x = cal.cities$long[top7], y = cal.cities$lat[top7], pch = 19)

## add a constant to latitude to avoid overlapping with circles
text(x = cal.cities$long[top7] + 2.25, y = cal.cities$lat[top7],
     label = cal.cities$name[top7])
title("Largest cities of California")
```

# Part 1:Map Map Map! YOUR TURN

What is your favorite state? Plot it and their top seven cities by following the code we just saw above!

```{r student }
map(database = "state", regions = "Connecticut")
```

```{r student2 }
ct.cities <- subset(us.cities, subset = (country.etc == "CT"))
sind <- order(ct.cities$pop, decreasing = TRUE) # order by population
top7 <- sind[1:7] # seven cities with largest population

map(database = "state", regions = "Connecticut")

points(x = ct.cities$long[top7], y = ct.cities$lat[top7], pch = 19)

text(x = ct.cities$long[top7] , y = ct.cities$lat[top7], label = ct.cities$name[top7])
title("Largest cities of Connecticut")

```

Good Jobs!!! Now you master how to use map() function and add title/text/ points etc. to decorate it !


# Part 2: Election!!!
Are you bored with just mapping some state? Let's do a bit more fun visualization with election vote data in US!
### Preparation
First, we will read the csv file from the following url.
```{r c }
pres08 <- read.csv(url("https://raw.githubusercontent.com/NorikaNarimatsu/tutorial_geospatial_analysis/main/pres08.csv"))
```

### California Vote 
Then, this is a recap from part 1. We will map the California and specify its color as blue because they are known as democratic state. 
```{r ca3 }
## California as a blue state
map(database = "state", regions = "California", col = "blue", 
    fill = TRUE)
```



Hmmm, is it still boring? Then, let's try to calculate the vote share and find the shade of purple!
The code creates a color for California based on the vote share. This is done by using the rgb() function to create a color that is a combination of red and blue, with the amount of red and blue determined by the Republican and Democratic vote shares, respectively.

```{r ca4 }
## two-party vote share
pres08$Dem <- pres08$Obama / (pres08$Obama + pres08$McCain)
pres08$Rep <- pres08$McCain / (pres08$Obama + pres08$McCain)

## color for California
cal.color <- rgb(red = pres08$Rep[pres08$state == "CA"], 
                 blue = pres08$Dem[pres08$state == "CA"], 
                 green = 0)

## California as a purple state
map(database = "state", regions = "California", col = cal.color,
    fill = TRUE)
```



Nice! Let's expand out analysis to whole USA! The code then creates a map of the entire United States, with each state colored either red or blue depending on whether the Republican or Democratic candidate won the state. The code uses a for loop to iterate over each row of the dataset and call the map() function for each state.

```{r usa_08 }

## America as red and blue states
map(database = "state") # for some reason this needs to be repeated twice to get the map correct if you run markdown
for (i in 1:nrow(pres08)) {
    if ((pres08$state[i] != "HI") & (pres08$state[i] != "AK") &
        (pres08$state[i] != "DC")) {
        maps::map(database = "state", regions = pres08$state.name[i],
            col = ifelse(pres08$Rep[i] > pres08$Dem[i], "red", "blue"),
            fill = TRUE, add = TRUE)
    }
}
title("US Election 2008") # add a title
```



Yes !!! Let's try the shade version too! 
```{r usa2_12 }

## America as purple states
map(database = "state") # create a map
for (i in 1:nrow(pres08)) {
    if ((pres08$state[i] != "HI") & (pres08$state[i] != "AK") &
        (pres08$state[i] != "DC")) {
        map(database = "state", regions = pres08$state.name[i], 
            col = rgb(red = pres08$Rep[i], blue = pres08$Dem[i],
                green = 0), fill = TRUE, add = TRUE)
    }
}
title("US Election 2008") # add a title
```
Perfecct! Now it is your turn to do the same analysis with data from 2012 election!

# Part 2: Election! YOUR TURN!!!
Since there is no column called "state.name" in pres12 dataset, we did data preparation for you! In the cells below, you should visualize the 1) the red and blue map for US, and 2) the shade of purple map for US! 

```{r president2 }

pres12 <- read.csv(url("https://raw.githubusercontent.com/NorikaNarimatsu/tutorial_geospatial_analysis/main/pres12.csv"))
# select the state and state.name columns from pres08
pres08_states <- pres08[, c("state", "state.name")]
# merge the state.name column from pres08 into pres12 based on the common state column
pres12 <- merge(pres12, pres08_states, by = "state", all.x = TRUE)
# view the merged dataframe
head(pres12)
```

### 1) the red and blue map for US,
EXTRA CELL
```{r usa_student1 }

## two-party vote share
pres12$Dem <- pres12$Obama / (pres12$Obama + pres12$Romney)
pres12$Rep <- pres12$Romney / (pres12$Obama + pres12$Romney)

# America as red and blue states
map(database = "state") # for some reason this needs to be repeated twice to get the map correct if you run markdown

for (i in 1:nrow(pres12)) {
    if ((pres12$state[i] != "HI") & (pres12$state[i] != "AK") &
        (pres12$state[i] != "DC")) {
        maps::map(database = "state", regions = pres12$state.name[i],
            col = ifelse(pres12$Rep[i] > pres12$Dem[i], "red", "blue"),
            fill = TRUE, add = TRUE)
    }
}

title("US Election 2012") # add a title
```
### 2) the shade of purple map for US! 
```{r usa_student2 }

## America as purple states
map(database = "state") # create a map
for (i in 1:nrow(pres12)) {
    if ((pres12$state[i] != "HI") & (pres12$state[i] != "AK") &
        (pres12$state[i] != "DC")) {
        map(database = "state", regions = pres12$state.name[i], 
            col = rgb(red = pres12$Rep[i], blue = pres12$Dem[i],
                green = 0), fill = TRUE, add = TRUE)
    }
}
title("US Election 2012") # add a title
```
### Questions 2:
Did you find any difference in the vote share between 2008 and 2012? 

Indiana: In 2008, Indiana voted for the Republican candidate John McCain, but in 2012, it voted for the Democratic candidate Barack Obama.

North Carolina: In 2008, North Carolina voted for Obama, but in 2012, it voted for the Republican candidate Mitt Romney.

```{r majority }
# Create a data frame to store the winning party in each state for the two elections
results <- data.frame(state = pres08$state, dem_08 =pres08$Dem , dem_12 = pres12$Dem,
                      winner08 = ifelse(pres08$Dem > 0.5, "Democrat", "Republican"),
                      winner12 = ifelse(pres12$Dem > 0.5, "Democrat", "Republican"))

# Find the states that switched from one party to the other between the two elections
switched <- subset(results, winner08 != winner12)

switched$vote_share_change <- switched$dem_08 - switched$dem_12

# Print the results
switched


```


### Questions 3:
Do you have any remaining questions about this pcw, especially coding part? Write down your questions here and bring your confusions to the class! 
EXTRA CELL


